---
title: "Nonparamatric_project"
author: "Haichuan Du"
date: "7/6/2020"
output:
  word_document: default
---


```{r}
library(tidyverse)
library(BSDA)
library(clinfun)
library(jmuOutlier) 
library(gmodels)
library(testforDEP)
library(ggplot2)
library(corrplot)
library(testforDEP) #Testing for dependency using Hoeffding's test and others.
```

data source:  https://www.kaggle.com/blastchar/telco-customer-churn
```{r}
setwd("~/Desktop/statistics/Nonparametric/project")
```
customerID: Customer ID
genderCustomer: gender (female, male)
SeniorCitizen: Whether the customer is a senior citizen or not (1, 0)
PartnerWhether: the customer has a partner or not (Yes, No)
Dependents: Whether the customer has dependents or not (Yes, No)
tenure: Number of months the customer has stayed with the company<-----------------------------------------
PhoneService: Whether the customer has a phone service or not (Yes, No)
MultipleLines: Whether the customer has multiple lines or not (Yes, No, No phone service)
InternetService: Customer’s internet service provider (DSL, Fiber optic, No)
OnlineSecurity: Whether the customer has online security or not (Yes, No, No internet service)
OnlineBackup: Whether the customer has online backup or not (Yes, No, No internet service)
DeviceProtection: Whether the customer has device protection or not (Yes, No, No internet service)
TechSupport: Whether the customer has tech support or not (Yes, No, No internet service)
StreamingTV: Whether the customer has streaming TV or not (Yes, No, No internet service)
StreamingMovies: Whether the customer has streaming movies or not (Yes, No, No internet service)
Contract: The contract term of the customer (Month-to-month, One year, Two year)
PaperlessBilling: Whether the customer has paperless billing or not (Yes, No)
PaymentMethod: The customer’s payment method (Electronic check, Mailed check, Bank transfer (automatic), Credit card (automatic))
MonthlyCharges: The amount charged to the customer monthly
TotalCharges: The total amount charged to the customer   <----------------------------------
Churn: Whether the customer churned or not (Yes or No) <-------------------------------------

# Read data
```{r}
churn_data <- read.csv("churn.csv")
churnT <- churn_data
head(churn_data)
attach(churn_data)
```
#Exploratory data analysis

Check dimension
```{r}
dim(churn_data)
```

check data format
```{r}
str(churn_data)
```

check  NA 
```{r}
num.NA <- sort(colSums(is.na(churn_data)), decreasing = TRUE)
subset(churn_data, is.na(churn_data$TotalCharges))
churn_data <- churn_data[-c(which(is.na(churn_data$TotalCharges))),]
```
Since the number of rows with NaN values are very small, and taking into consideration that all of them are churn=”NO”, which is over-represented in the data. We will just delete those rows from the working data


```{r}
# create a vertors changing tenure from monthly to year
churn_data$tenureToYear <- cut(churn_data$tenure, c(0,12,24,36,48,60,72), labels=c(1:6) )
churn_data$SeniorCitizen <- as.factor(churn_data$SeniorCitizen)
# Create vectors of Yes_churn and No_churn
Yes_churn <- churn_data %>%
  filter(Churn == "Yes")
No_churn <- churn_data %>%
  filter(Churn == "No")

Yes_Partner <- churn_data %>%
  filter(Partner == "Yes")
No_Partner <- churn_data %>%
  filter(Partner == "No")
```


```{r}
library(miscset)
#user attributes
ggplotGrid(ncol=2,lapply(c("gender","SeniorCitizen","Partner","Dependents"),
           function(col){ggplot(churn_data,aes_string(col)) + geom_bar(aes(fill=Churn),position="dodge")}))

#contract attributes
ggplotGrid(ncol=2,lapply(c("Contract","tenureToYear","PaperlessBilling", "PaymentMethod"),
           function(col){ggplot(churn_data,aes_string(col)) + geom_bar(aes(fill=Churn),position="dodge")}))

ggplotGrid(ncol=2,lapply(c("tenure", "MonthlyCharges","TotalCharges"),
           function(col){ggplot(churn_data,aes_string(col)) + geom_density(aes(col=Churn))}))

#service attribures
ggplotGrid(ncol=2,lapply(c("PhoneService","MultipleLines","InternetService"),
           function(col){ggplot(churn_data,aes_string(col)) + geom_bar(aes(fill=Churn),position="dodge")}))

ggplotGrid(ncol=2,lapply(c("TechSupport","OnlineSecurity","OnlineBackup","DeviceProtection","StreamingTV","StreamingMovies"),
           function(col){ggplot(churn_data,aes_string(col)) + geom_bar(aes(fill=Churn),position="dodge")}))

```

Then I try both parametric and nonparametric test to see if churn is associated with these variables.
```{r}
CrossTable(Churn, gender, chisq = T, format = "SPSS")                 # p =  0.4698339
chisq.test(Churn, gender, simulate.p.value=TRUE)                      # p-value = 0.4838

CrossTable(Churn, SeniorCitizen, chisq = T, format = "SPSS")          # p =  9.477904e-3
chisq.test(Churn, SeniorCitizen, simulate.p.value=T)                  # p-value = 0.0004998
#Partner
CrossTable(Churn, Partner, chisq = T, format = "SPSS")                # p =  1.519037e-36. 
chisq.test(Churn, Partner, simulate.p.value=T)                        # p-value = 0.0004998.
# Dependents
CrossTable(Churn, Dependents, chisq = T, format = "SPSS")             # p =  3.276083e-43.  
chisq.test(Churn, Dependents, simulate.p.value=T)                     # p-value = 0.0004998

```

K-S Test: (Monthly Charges)
```{r}
#kernal density plot
plot(density(Yes_churn$MonthlyCharges), col = 2, main = "Density plot of MonthlyCharge", xlab = "MonthlyCharge")
lines(density(No_churn$MonthlyCharges), col = 3)
  legend("topright", c("Yes_Churn", "No_Churn"),
         col = c(2, 3), lty = 1, cex=0.8, title = "Monthlycharge for Churn")

#boxplot
ggplot(data = churn_data, aes(Churn, MonthlyCharges, fill = Churn)) +
  geom_boxplot(outlier.colour = "blue") +
  labs(title = "Box plot of MonthlyCharges against Churn") +
  theme(plot.title = element_text(hjus = 0.5)) 

ks.test(Yes_churn$MonthlyCharges, No_churn$MonthlyCharges)

plot(c(0,sort(Yes_churn$MonthlyCharges),200),c(0,1:length(Yes_churn$MonthlyCharges),length(Yes_churn$MonthlyCharges))/length(Yes_churn$MonthlyCharges),xlab='MonthlyCharge',type='s',ylab='EDF')
lines(c(0,sort(No_churn$MonthlyCharges),200),c(0,1:length(No_churn$MonthlyCharges),length(No_churn$MonthlyCharges))/length(No_churn$MonthlyCharges),type='s',lty=2)

# I also use sign test to see If my data match up with national averages 65.5 for monthly charge ?
SIGN.test(MonthlyCharges, md = 65.5, conf.level = 0.95)
t.test(MonthlyCharges, mu = 65.5, conf.level = 0.95)            
```

Wilcoxon Rank Sum Test: (Monthly Charges by Partner)
```{r}
ggplot(data = churn_data, aes(Partner, MonthlyCharges, fill = Partner)) +
  geom_boxplot(outlier.colour = "blue") +
  labs(title = "Box plot of MonthlyCharges against Partner") +
  theme(plot.title = element_text(hjus = 0.5)) 

wilcox.test(Yes_Partner$MonthlyCharges, No_Partner$MonthlyCharges, conf.int = T)
t.test(Yes_Partner$MonthlyCharges, No_Partner$MonthlyCharges)
```

Association: Monthlycharges Vs. Tenure
```{r}
ggplot(data = churn_data, aes(tenure, MonthlyCharges, color = Contract)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "MonthlyCharges Vs. Tenure") +
  xlab("Tenure (months)") +
  theme(plot.title = element_text(hjus = 0.5))

ggplot(data = churn_data, aes(tenure, MonthlyCharges, color = Churn)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "MonthlyCharges Vs. Tenure") +
  xlab("Tenure (months)") +
  theme(plot.title = element_text(hjus = 0.5))

corrplot(cor(churn_data[sapply(churn_data, is.numeric)]), order="hclust", method = "number")
perm.cor.test(tenure, MonthlyCharges, method='spearman') 
cor.test(tenure, MonthlyCharges,method='pearson')
```

Does the contract have any effect on customer churn?
```{r}
ggplot(churn_data, aes(x = Contract, fill = Churn)) +
  geom_bar() +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3) +
  labs(title="Churn rate by contract status")
# The churn rate of month-to-month contract customers is much higher than the longer contract customers. Customers who are more willing to commit to longer contracts are less likely to leave.

ggplot(data = churn_data, aes(Churn, MonthlyCharges, fill = Contract)) +
  geom_boxplot(outlier.colour = "blue") +
  labs(title = "Box plot of MonthlyCharges for Chutract and Churn") +
  theme(plot.title = element_text(hjus = 0.5)) 

# plot(density(subset(No_churn, Contract=="One year")$MonthlyCharges), col = 2)
# lines(density(subset(No_churn, Contract=="Month-to-month")$MonthlyCharges), col = 3) 
# lines((density(subset(No_churn, Contract=="Two year")$MonthlyCharges)))

#Ansari-Bratest Test: One Year  Vs.  Month-to-month (Is there much difference between a monthly contract and an annual contract?)
ansari.test(subset(No_churn, Contract=="One year")$MonthlyCharges, subset(No_churn, Contract=="Month-to-month")$MonthlyCharges, alternative = "greater") 
#Running the normal-theory F test.
var.test(subset(No_churn, Contract=="One year")$MonthlyCharges, subset(No_churn, Contract=="Month-to-month")$MonthlyCharges, alternative = "greater") 

#Wilcoxon Rank Sum Test: (Two Year Vs. One Year)  Is the monthly charge for one year and two years the same?
wilcox.test(subset(No_churn, Contract=="One year")$MonthlyCharges, subset(No_churn, Contract=="Two year")$MonthlyCharges, conf.int = T)  
t.test(subset(No_churn, Contract=="One year")$MonthlyCharges, subset(No_churn, Contract=="Two year")$MonthlyCharges) #p < 0.05 

# median(subset(No_churn, Contract=="One year")$MonthlyCharges)

#Kruskal–Wallis Test: (Month-to-month  Vs. Two Year Vs. One Year)  Is the monthly charge the same for all different contract?
data_Monthlycharge <- c(subset(No_churn, Contract=="Month-to-month")$MonthlyCharges,
                 subset(No_churn, Contract=="One year")$MonthlyCharges,
                 subset(No_churn, Contract=="Two year")$MonthlyCharges) #Specifying which values are for which samples.
groups=rep(1:3,c(2220,1306,1637)) #Specifying which values are for which samples.

kruskal.test(data_Monthlycharge, groups)   # p value = 0.32
summary(aov(data_Monthlycharge~factor(groups)))

#Now develop the nonparametric approach.
#First simulate to find the appropriate 
# level for the pairwise intervals.
#Second, compute the pairwise intervals 
# for the earlier set of data.

simultaneous8452=function(data,groups,conf.level){
  #Note: The groups must be numbered from 1 to k.
  k=max(groups) #Finding the number of groups.
  nvec=double(k)
  for (i in 1:k){
    nvec[i]=sum(groups==i)} #Finding the sample sizes.
  nruns=10000
  minalpha=double(nruns)
  N=sum(nvec) #Total sample size.
  for (run in 1:nruns){
    minalpha[run]=1
    simdata=rnorm(N)
    for (i in 1:(k-1)){
      for (j in (i+1):k){
        minalpha[run]=min(minalpha[run],wilcox.test(simdata[groups==i],simdata[groups==j])$p.value)}}}
  crit2=quantile(minalpha,1-conf.level)
  crit=max(minalpha[minalpha<crit2/1.000001])
  cov=1-mean(minalpha<=crit*1.000001)
  
  #Now finding the conservative intervals.
  ct=0
  for (i in 1:(k-1)){
    for (j in (i+1):k){
      if (ct==0){mat=c(i,j,wilcox.test(data[groups==i],data[groups==j],conf.int=T,dcl=(1-crit)*1.000001)$conf.int)}
      if (ct>0){mat=rbind(mat,c(i,j,wilcox.test(data[groups==i],data[groups==j],conf.int=T,dcl=(1-crit)*1.000001)$conf.int))}
      ct=ct+1}}
  mat
  
  list(crit=crit,cov=cov,mat=mat)}
simultaneous8452(data_Monthlycharge,groups,0.95)

summary(aov(data_Monthlycharge~factor(groups))) #anova F test P = 0.1
TukeyHSD(aov(data_Monthlycharge~factor(groups)))  #Now finding Tukey's pairwise CIs 

# 
# data_Monthlycharge2 <- c(subset(Yes_churn, Contract=="Month-to-month")$MonthlyCharges,
#                  subset(Yes_churn, Contract=="One year")$MonthlyCharges,
#                  subset(Yes_churn, Contract=="Two year")$MonthlyCharges)
# 
# groups2=rep(1:3,c(1655,166,48)) #Specifying which values are for which samples.
# 
# jonckheere.test(data_Monthlycharge2, groups2, nperm = 2000, alternative = "increasing") #Jonckheere-Terpstra test. (clinfun)
```


Chi-Square Test/permutation chi square for churn vs service attributes
```{r}
# PhoneService
CrossTable(Churn, PhoneService, chisq = T, format = "SPSS")
chisq.test(Churn, PhoneService, simulate.p.value=T)
# MultipleLines
CrossTable(Churn, MultipleLines, chisq = T, format = "SPSS")
chisq.test(Churn, MultipleLines, simulate.p.value=T)
# OnlineSecurity
CrossTable(Churn, OnlineSecurity, chisq = T, format = "SPSS")
chisq.test(Churn, OnlineSecurity, simulate.p.value=T)
# OnlineBackup
CrossTable(Churn, OnlineBackup, chisq = T, format = "SPSS")
chisq.test(Churn, OnlineBackup, simulate.p.value=T)
# DeviceProtection
CrossTable(Churn, DeviceProtection, chisq = T, format = "SPSS")
chisq.test(Churn, DeviceProtection, simulate.p.value=T)
# TechSupport
CrossTable(Churn, TechSupport, chisq = T, format = "SPSS")
chisq.test(Churn, TechSupport, simulate.p.value=T)
# StreamingMovies
CrossTable(Churn, StreamingMovies, chisq = T, format = "SPSS")
chisq.test(Churn, StreamingMovies, simulate.p.value=T)
# StreamingTV
CrossTable(Churn, StreamingTV, chisq = T, format = "SPSS")
chisq.test(Churn, StreamingTV, simulate.p.value=T)
# Contract
CrossTable(Churn, Contract, chisq = T, format = "SPSS")
chisq.test(Churn, Contract, simulate.p.value=T)
```

Bootstrap
```{r}
oursample = churn_data
index = nrow(oursample)
Tenure_thetahat <- glm(Churn~tenure+MonthlyCharges+PhoneService+InternetService, family = "binomial", data = oursample)$coeff[2]
MonthlyCharges_thetahat <- glm(Churn~tenure+MonthlyCharges+PhoneService+InternetService, family = "binomial", data = oursample)$coeff[3]
PhoneService_thetahat <- glm(Churn~tenure+MonthlyCharges+PhoneService+InternetService, family = "binomial", data = oursample)$coeff[4]
InternetService_thetahat <- glm(Churn~tenure+MonthlyCharges+PhoneService+InternetService, family = "binomial", data = oursample)$coeff[5]

Tenure_thetahat
MonthlyCharges_thetahat
PhoneService_thetahat
InternetService_thetahat

Tenure_thetahat.b = rep(NA,1000); MonthlyCharges_thetahat.b = rep(NA,1000); PhoneService_thetahat.b = rep(NA,1000); InternetService_thetahat.b = rep(NA,1000)
for (i in 1:1000) {
  ### draw the bootstrap sample and calculate thetahat.b
  bootindex = sample(1:index, index, replace=T)
  bootsample = oursample[bootindex,]
  Tenure_thetahat.b[i] = glm(Churn~tenure+MonthlyCharges+PhoneService+InternetService, family = "binomial", data = bootsample)$coeff[2]
  MonthlyCharges_thetahat.b[i] = glm(Churn~tenure+MonthlyCharges+PhoneService+InternetService, family = "binomial", data = bootsample)$coeff[3]
  PhoneService_thetahat.b[i] = glm(Churn~tenure+MonthlyCharges+PhoneService+InternetService, family = "binomial", data = bootsample)$coeff[4]
  InternetService_thetahat.b[i] = glm(Churn~tenure+MonthlyCharges+PhoneService+InternetService, family = "binomial", data = bootsample)$coeff[5]
}

par(mfrow=c(2,2))
hist(Tenure_thetahat.b); hist(MonthlyCharges_thetahat.b); hist(PhoneService_thetahat.b); hist(InternetService_thetahat.b)

quantile(Tenure_thetahat.b,c(0.025,0.975))
quantile(MonthlyCharges_thetahat.b,c(0.025,0.975))
quantile(PhoneService_thetahat.b,c(0.025,0.975))
quantile(InternetService_thetahat.b,c(0.025,0.975))
```



